---
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, cache = T, warning = F)
```

# Cross-Scale Aggregation (CSA) framework

## Spatiotemporal applications over the Netherlands 

### Introduction

In this introductory presentation, the CSA methodology is applied for the comparison of different data sets. This includes observational and model simulation data products from various sources, i.e., the satellite GPM IMERG data product, KNMI station and radar data sets, as well as, CNRM simulation and NCEP/NCAR reanalysis. Two examples are given, one in temporal and one in spatial domain. In the first case, the aggregation curves of the dataset are estimated and then plotted together. In the second application, the two-dimensional aggregation curves of six  precipitation events which have similar mean over the whole domain are compared. 

The full description of the methodolog and the examples can be found in the journal of [Geoscientific Model Development](https://www.geoscientific-model-development.net/): 

Markonis et al., Cross-scale analysis framework for model/data comparison and integration, Geoscientific Model Development, _under review_.

***

### Functions

The csa package can be downloaded from github with package devtools:

```{r, eval = F}
devtools::install_github("imarkonis/csa")
```

**csa** Computes (and by default plots) the aggregation curve of a given statistic in a single dimension, e.g., time. The statistics, that can be estimated across the cross-scale continuum, are: variance ("var"; default), standard deviation ("sd"), skewness ("skew"), kurtosis ("kurt"), L-scale ("l2"), coefficient of L-variation ("t2"), L-skewness ("t3"), L-kurtosis ("t4").

**csas** As above but in two dimensions, i.e., space. Input variable is in brick format (raster), while output is in tidy format (data.table), with each layer of the initial brick coresponding to a different variable number. This allows for further extending CSA plots to both space and time.

**csa.plot** Plots a data.table in the CSA format (one column for scale number and one for statistic value). If _wn_ argument is true, then the aggregation curve of white noise process is also plotted. 

**csa.multiplot** As above for a group of aggregation curves. In the input data.table, a column describing the data set is added, while the argument _smooth_ allows for loess smoothing of the aggregation curves. If there are more than 10 lines then no legend is ploted and the lines are transparent.

**csa.rescale** Auxiliary function for matching and comparing aggregation curves estimated at different scales.

**dt.to.brick** Auxiliary function for transforming data table objects to brick format, for integration in _csas_ function.

***

More can be found in the accompanying help files of each function, as well as in two interactive applications about [aggregation](https://markonis.shinyapps.io/shiny_agg/) and the aggregation curves of some well-known [stochastic models](https://markonis.shinyapps.io/shiny_csa/).

***

### Data

```{r data, echo = F}
load("../data/observ_raw.rdata") 
load("../data/model_raw.rdata") 
```

**Satellite data (GPM)**
  
* Dataset: GPM IMERG Final Precipitation 
* Temporal Resolution: 24 hours
* Spatial Resolution: 0.1 x 0.1 deg
* Date Range: `r min(gpm_prcp$time)` to `r max(gpm_prcp$time)` 
* Spatial Region: latitude: `r range(gpm_cells$lat)`, longitude: `r range(gpm_cells$lon)`
* Format: netCDF 4
* Variables: gpm_cells (coordinates) & gpm_prcp (values) in _observ_raw.Rdata_
* IDs: `r min(gpm_cells$id)`
* Size: `r nrow(gpm_prcp)` total values, multiplied by `r nrow(gpm_cells)` grid cells
* Downloaded from: https://climexp.knmi.nl/select.cgi?id=312456c83e660703df1bfea9ba4fba50&field=imerg_daily 

**Station data (KNMI)**
  
* Dataset: 240 homogenised stations 1951-now
* Temporal Resolution: 24 hours
* Spatial Resolution: -
* Date Range: `r min(knmi_prcp$time)` to `r max(knmi_prcp$time)` 
* Spatial Region: latitude: `r range(knmi_stations$lat)`, longitude: `r range(knmi_stations$lon)`
* Format: Text file
* Variables: knmi_stations (coordinates) & knmi_prcp (values) in _observ_raw.Rdata_
* IDs: `r min(knmi_stations$id)`
* Size: `r nrow(knmi_prcp)` total values, multiplied by `r nrow(knmi_stations)` stations
* Downloaded from: https://climexp.knmi.nl/PhomNL.cgi?id=312456c83e660703df1bfea9ba4fba50

**Radar data (KNMI)**
  
* Dataset: RAD_NL25_RAC_MFBS_24H_NC
* Temporal Resolution: 24 hours
* Spatial Resolution: 1 x 1 km
* Date Range: `r min(rdr_prcp$time)` to `r max(rdr_prcp$time)` 
* Spatial Region: latitude: `r range(rdr_cells$lat)`, longitude: `r range(rdr_cells$lon)`
* Format: netCDF 4
* Variables: rdr_cells (coordinates) & rdr_prcp (values) in _observ_raw.Rdata_
* IDs: `r min(rdr_cells$id)`
* Size: `r nrow(rdr_prcp)` total values, multiplied by `r nrow(rdr_cells)` grid cells
* Downloaded from: https://climexp.knmi.nl/select.cgi?id=312456c83e660703df1bfea9ba4fba50&field=knmi_radar_daily

  **Reanalysis data (NCEP/NCAR)**
  
* Dataset: NMC reanalysis 
* Temporal Resolution: 24 hours
* Spatial Resolution: 2.5 x 2.5 deg 
* Date Range: `r min(ncep_prcp$time)` to `r max(ncep_prcp$time)` 
* Spatial Region: 1 grid cell at 52.38N & 5.625E 
* Format: netCDF 4
* Variables: ncep_prcp (values) in _model_raw.Rdata_
* Size: `r nrow(ncep_prcp)` total values in 1 grid cell
* Downloaded from: https://climexp.knmi.nl/select.cgi?id=312456c83e660703df1bfea9ba4fba50&field=nprate_daily

**Simulation data (CNRM)**
  
* Dataset: cnrm cm3 20c3m pr
* Temporal Resolution: 24 hours
* Spatial Resolution: 2.8 x 2.8 deg 
* Date Range: `r min(cnrm_prcp$time)` to `r max(cnrm_prcp$time)` 
* Spatial Region: 1 grid cell at 51.625N & 5.625E 
* Format: netCDF 4
* Variables: cnrm_prcp (values) in _model_raw.Rdata_
* Size: `r nrow(cnrm_prcp)` total values in 1 grid cell
* Downloaded from: https://climexp.knmi.nl/select.cgi?id=312456c83e660703df1bfea9ba4fba50&field=pr_A2_cnrm_cm3_20c3m

***
  
### Code structure

The code needed to reproduce the importing and preparation of the datasets provided by the csa package consists of two R scripts. These are:

[data_import.R](https://github.com/imarkonis/CrossScale/blob/master/bin/data_import.R) Imports observed and modeled data for the Netherlands. Observational data are composed of two data tables, one for values and one for cell/station coordinates and then saved in _observ_raw.Rdata_. Model simulation data represent a single grid cell (coarser resolution) and thus a single data table is stored in _model_raw.Rdata_.

[data_prep.R](https://github.com/imarkonis/CrossScale/blob/master/bin/data_prep.R) Prepares datasets for comparison. This script reads the files prepared in _data_import.R_ and makes the necessary changes for the application of csa in time and space. For the temporal domain, all observational datasets are aggregated (average) over the Netherlands and then saved in _temporal.Rdata_. For the spatial domain, 6 precipitation events (daily) with mean precipitation between 10 and 11 mm, are randomly picked and then stored in _spatial.Rdata_. 

Although here, we will use directly the data from the csa package, we have included the links for reasons of reproducibility. 

```{r data load}
data(gpm_nl, knmi_nl, rdr_nl, ncep_nl, cnrm_nl, gpm_events)
```

The libraries needed are:

```{r libraries & data, include = F}
library(data.table); library(reshape2); library(csa); library(raster)
```

```{r libraries, eval = F}
library(data.table); library(reshape2); library(csa); library(raster)
```

The examples presented below can also be found in:

[temporal_agg.R](https://github.com/imarkonis/CrossScale/blob/master/bin/spatial_agg.R) A demonstration of csa function in temporal domain for variance, as well as csa.multiplot for the 5 datasets.

[spatial_agg.R](https://github.com/imarkonis/CrossScale/blob/master/bin/temporal_agg.R) Same as above, but for GPM dataset in spatial domain.

***

### Results

#### Temporal Aggregation

To estimate and plot the aggregation curve of variance (function default), the _csa_ function is used with a time series in vector format as input: 

```{r csa_1, eval = F}
csa(gpm_nl$prcp)
```

```{r csa_2, fig.width = 4, fig.height = 3, echo = F, fig.align = 'center'}
res <- csa(gpm_nl$prcp)$plot
```

To compare the aggregation curves of different data sets, the _csa.multiplot_ function can be applied. Below three different utilizations of the function are presented (file _temporal_agg.R_). In the first case, the common arguments for the CSA plot are used, while in the second only the logarithmic scales are used (argument _fast = T_), which could be handy if speeding up the calculations is needed. The third plot, uses the csa.multiplot argument _smooth_ to produce a loess smoothed line for each aggregation curve, as well as the _wn_ argument to compare with the white noise process.

```{r temporal_agg, fig.align = 'center'}
set_1 <- data.frame(csa(gpm_nl$prcp, plot = F), dataset = "gpm")
set_2 <- data.frame(csa(rdr_nl$prcp, plot = F), dataset = "radar")
set_3 <- data.frame(csa(knmi_nl$prcp, plot = F), dataset = "station")
set_4 <- data.frame(csa(ncep_nl$prcp, plot = F), dataset = "ncep")
set_5 <- data.frame(csa(cnrm_nl$prcp, plot = F), dataset = "cnrm")

g1 <- csa.multiplot(rbind(set_1, set_2, set_3, set_4, set_5))
print(g1)

set_1 <- data.frame(csa(gpm_nl$prcp, plot = F, fast = T), dataset = "gpm")
set_2 <- data.frame(csa(rdr_nl$prcp, plot = F, fast = T), dataset = "radar")
set_3 <- data.frame(csa(knmi_nl$prcp, plot = F, fast = T), dataset = "station")
set_4 <- data.frame(csa(ncep_nl$prcp, plot = F, fast = T), dataset = "ncep")
set_5 <- data.frame(csa(cnrm_nl$prcp, plot = F, fast = T), dataset = "cnrm")

g2 <- csa.multiplot(rbind(set_1, set_2, set_3, set_4, set_5))
print(g2)

g3 <- csa.multiplot(rbind(set_1, set_2, set_3, set_4, set_5), smooth = T, wn = T)
print(g3)
```

We can also use the _csa.multiplot_ to represent the variance ratio between one dataset (here knmi stations) with the others. 

```{r scaling comparison, fig.align = 'center'}
set_1_comp <- set_3[set_3$scale %in% set_1$scale, ]
set_1_comp$dataset <- set_1$dataset
set_1_comp$var <- set_1_comp$var/set_1$var

set_2_comp <- set_3[set_3$scale %in% set_2$scale, ]
set_2_comp$dataset <- set_2$dataset
set_2_comp$var  <- set_2_comp$var/set_2$var

set_4_comp <- set_3[set_3$scale %in% set_4$scale, ]
set_4_comp$dataset <- set_4$dataset
set_4_comp$var <- set_4_comp$var/set_4$var

set_5_comp <- set_3[set_3$scale %in% set_5$scale, ]
set_5_comp$dataset <- set_5$dataset
set_5_comp$var <- set_5_comp$var/set_5$var

set_comp <- rbind(set_1_comp, set_2_comp, set_4_comp, set_5_comp)
csa.multiplot(set_comp, log_x = T, log_y = F)
```

We can see that the model results are closer to station until the 10-day scale and then begin to overestimate the variability, whereas the opposite behaviour is observed from scale 2, in the remote sensing data (satellite and radar).

***

#### Spatial Aggregation

For the application in the spatial domain, six daily precipitation events that share average precipitation over the Netherlands (between 10 and 11 mm) are selected:

```{r spatial_agg_1, fig.align = 'center'}
event_dates <- format(gpm_events[, unique(time)], "%d-%m-%Y") 
gpm_events_brick <- dt.to.brick(gpm_events, var_name = "prcp")
plot(gpm_events_brick, col = rev(colorspace::sequential_hcl(40)),
     main = event_dates)
```

The application of _csas_ is quite similar to _csa_: 

```{r spatial_agg_2, fig.align = 'center'}
g4 <-csas(gpm_events_brick)
g4$plot

gpm_sp_scale <- csas(gpm_events_brick, plot = F)
gpm_sp_scale[, variable := factor(variable, labels = event_dates)]
csa.multiplot(gpm_sp_scale, smooth = T, log_x = F, log_y = F) 
```

Here, we see how the variance in space changes as the scale gets coarser (to a threshold of 30 grid cells as a final scale).

***

### Further Reading

Interested readers might find useful

***








